---
title: "MicrobiomeBenchmarkData"
author:
  - name: "Samuel D. Gamboa-Tuz"
    email: "Samuel.Gamboa.Tuz@gmail.com"
  - name: "Levi Waldron"
package: MicrobiomeBenchmarkData
output: 
  BiocStyle::html_document:
    toc: true
abstract: >
  Document summary
vignette: >
  %\VignetteIndexEntry{MicrobiomeBenchmarkData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `MicrobiomeBenchamrkData` package provides a collection of datasets suitable
for benchmarking of differential abundance methods in R.

The datasets are also available through
[Zenodo (sandbox)](https://sandbox.zenodo.org/record/946131) so they can be
downloaded and used out of the R environment if preferred/needed.

The main characteristic of this collection of datasets is that they contain
some biological truth that has been proved experimentally in the literature.

# Installation

MicrobiomeBenchmarkData should be available through the BioConductor project.

```{r installation, eval=FALSE}
# if (!"BiocManager" %in% installed.packages()[,"Package"])
# install.packages("BiocManager")

## Release version
# BiocManager::install("MicrobiomeBenchmarkData")

## Development version
# BiocManager::install("waldronlab/MicrobiomeBenchmarkData", build_vignettes = TRUE)
```

```{r, message=FALSE}
library(MicrobiomeBenchmarkData)
library(purrr)
```

# Sample metadata

All sample metadata is merged into a single data frame and provided as data
object:

```{r}
data('sampleMetadata')
## Get columns present in all samples
sample_metadata <- sampleMetadata %>% 
    discard(~any(is.na(.x))) %>% 
    head()
knitr::kable(sample_metadata)
```

# Accessing datasets

Currently, there are `r length(MicrobiomeBenchmarkData::getDataset())`
datasets available through the MicrobiomeBenchmarkData. These datasets are
accessed through the `getDataset` function.

## Print avaialable datasets

If no arguments are provided, the list of available datasets is printed on
screen and returned invisiby as a character vector:

```{r}
dats <- getDataset()
```
```{r}
dats
```

## Access a single dataset

In order to import a dataset, the `getDataset` function must be used with the
name of the dataset as the first argument (`x`) and the `dryrun` argument set
to `FALSE`. The output is a list vector with the dataset imported as a
TreeSummarizedExperiment object.

```{r}
tse <- getDataset('HMP_2012_16S_gingival_V35_subset', dryrun = FALSE)[[1]]
tse
```

## Access a few datasets

Several datasets can be imported simultaneously by giving the names of the 
different datasets in a character vector:

```{r}
list_tse <- getDataset(dats[2:4], dryrun = FALSE)
str(list_tse, max.level = 1)
```

## Access all of the datasets

If all of the datasets must to be imported, this can be done by providing
the `dryrun = FALSE` argument alone.

```{r}
mbd <- getDataset(dryrun = FALSE)
str(mbd, max.level = 1)
```

# Annotations for each taxa are included in rowData

The biological annotations of each taxa are provided as a column in the
`rowData` slot of the TreeSummarizedExperiment.

```{r}
## In the case, the column is named as taxon_annotation 
tse <- mbd$HMP_2012_16S_gingival_V35_subset
rowData(tse)
```

# Cache 

The datasets are cached so they're only downloaded once. The cache and all of
the files contained in it can be removed with the `removeCache` function.

```{r, eval=FALSE}
# removeCache()
```


# Session information

```{r}
sessionInfo()
```

