---
title: "MicrobiomeBenchmarkData"
author:
  - name: "Samuel D. Gamboa-Tuz"
    email: "Samuel.Gamboa.Tuz@gmail.com"
  - name: "Levi Waldron"
package: MicrobiomeBenchmarkData
output: 
  BiocStyle::html_document:
    toc: true
abstract: >
  Document summary
vignette: >
  %\VignetteIndexEntry{MicrobiomeBenchmarkData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `MicrobiomeBenchamrkData` package provides a collection of datasets suitable
for benchmarking of differential abundance methods in R.

The datasets are also available through
[Zenodo (sandbox)](https://sandbox.zenodo.org/record/946131) so they can be
downloaded and used out of the R environment if preferred/needed.

The main characteristic of this collection of datasets is that they contain
some biological truth that has been proved experimentally in the literature.

# Installation

MicrobiomeBenchmarkData should be available through the BioConductor project.

```{r installation, eval=FALSE}
# if (!"BiocManager" %in% installed.packages()[,"Package"])
# install.packages("BiocManager")

## Release version
# BiocManager::install("MicrobiomeBenchmarkData")

## Development version
# BiocManager::install("waldronlab/MicrobiomeBenchmarkData")
```

```{r, message=FALSE}
library(MicrobiomeBenchmarkData)
```

# Accessing datasets

Datasets are accessed through the `getDataset` function.

## Print avaialable datasets


```{r}
dats <- getDataset()
```
```{r}
dats
```

## Access a single dataset

```{r}
smoking <- getDataset(dats[1], dryrun = FALSE)[[1]]
smoking
```

## Access a few datasets

```{r}
gingival <- getDataset(dats[2:5], dryrun = FALSE)
str(gingival, max.level = 1)
```

## Access all of the datasets

```{r}
mbd <- getDataset(dryrun = FALSE)
str(mbd, max.level = 1)
```

# Annotations for each taxa are included in rowData

Currently, only in "HMP_2012_16S_gingival_V35_subset".

```{r}
rowData(mbd$HMP_2012_16S_gingival_V35_subset)
```

# Caching

The datasets are cached so they're only downloaded once. The cache and all of
it's files can be removed with the `removeCache` function.

```{r, eval=FALSE}
# removeCache()
```


# Convert to phyloseq

The mia package can be used to convert a TreeSummarizedExperiment to a phyloseq
object, and use the dataset with other pipelines.

```{r}
ps <- mia::makePhyloseqFromTreeSummarizedExperiment(
    mbd$HMP_2012_16S_gingival_V35_subset
    )
ps
```




# Session information

```{r}
sessionInfo()
```

