---
title: "Recalibration of the count matrix of the Stammler_2016_16S_spikein dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{recalibrare_spikein_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

One of the main characteristics of the Stammler_2016_16S_spikein
dataset is the presence of spike-in bacteria with a known fixed amount of 
bacterial cells. These known loads of bacteria can be used to recalibrate the
raw counts of the matrix and obtain recalibrated absolute counts. In this 
vignette, we provide an example of how to recalibrate the counts of the count
matrix based on the read counts of *Salinibacter ruber*. This procedure is
referred to as  Spike-in-based calibration to total microbial load (SCML) in 
[Sammler et al., 2016](https://doi.org/10.1186/s40168-016-0175-0).

```{r setup, message=FALSE}
library(MicrobiomeBenchmarkData)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Import data

```{r}
tse <- getDataset('Stammler_2016_16S_spikein', dryrun = FALSE)[[1]]
counts <- assay(tse)
```

## Ids of the spike-in bacteria

Identifiers of the spiked-in bacteria have the suffix 'XXXX'.

| Bacteria | ID | Load |
| -------- | -- | ---- |
| *Salinibacter ruber* | AF323500XXXX | 3.0 x 10<sup>8</sup> |
| *Rhizobium radiobacter* | AB247615XXXX | 5.0 x 10<sup>8</sup> |
| *Alicyclobacillus acidiphilus* | AB076660XXXX | 1.0 x 10<sup>8</sup> |


## Recalibrate based on *Salinibacter ruber* abundance.

This recalibration is based on the original article. The only difference is that
the numbers have been rounded up to obtain counts.

```{r}
## AF323500XXXX is the unique OTU corresponding to S. ruber
s_ruber <- counts['AF323500XXXX', ]
size_factor <- s_ruber/mean(s_ruber)

SCML_data <- counts 
for(i in seq(ncol(SCML_data))){
    SCML_data[,i] <- round(SCML_data[,i] / size_factor[i])
}
```


Brief comparison of counts

```{r, fig.width=7}

no_cal <- counts %>% 
    colSums() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column(var = 'sample_id') %>% 
    magrittr::set_colnames(c('sample_id', 'colSum')) %>% 
    mutate(calibrated = 'no') %>% 
    as_tibble()

cal <-  SCML_data %>% 
    colSums() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column(var = 'sample_id') %>% 
    magrittr::set_colnames(c('sample_id', 'colSum')) %>% 
    mutate(calibrated = 'yes') %>% 
    as_tibble()

data <- bind_rows(no_cal, cal)

data %>% 
    ggplot(aes(sample_id, colSum)) + 
    geom_col(aes(fill = calibrated), position = 'dodge') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

The counts matrix can be replaced in the original tse in order to preserve the
same metadata.

```{r}
assay(tse) <- SCML_data
tse
```






